---
title: 'KAWN RAPID: Sensor Data Analysis'
author: "Michelle Catherine Kelly"
date: "3 June 2018"
output:
  html_document:
    theme: yeti
    df_print: paged
    toc: yes
    number_sections: true
---

```{r setup, echo = FALSE, warning=FALSE, error=FALSE, message=FALSE}
### Global options ###
knitr::opts_chunk$set(fig.height = 8, 
                      fig.width = 10, 
                      cache=TRUE) # enable caching globally

### General packages ###
library(XML)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(grid)
library(dplyr)
library(magrittr)
library(zoo)
library(wesanderson) # pallets
library(chron)
library(data.table)

### USGS loading ###
library(dataRetrieval)

### NEON loading ###
# library(devtools)
# install_github("NEONScience/NEON-utilities/neonUtilities", dependencies = TRUE)
library(neonUtilities)

### Stream metabolism ###
# install.packages("streamMetabolizer", dependencies = TRUE, repos = c("https://owi.usgs.gov/R", "https://cran.rstudio.com"))
library(streamMetabolizer)
library(rstan)
# install.packages("installr")
# library("installr")
# install.Rtools()

# check if Rtools is installed properly, output should be 4
# install.packages("Rcpp")
# library(Rcpp)
# Rcpp::evalCpp("2+2")
```

***  

# Overview

This is a **working document** for analyzing sensor data from the NSF RAPID Kansas River Nitrogen Cycling Project

***  

# Data Import  

This section contains all data loading and cleanup chunks  

Data pull = data is taken from an online host (USGS, NEON)
Data load = data is taken from local file

## Nitratax Data  

**Nitratax_XMLtoCSV:** Function to convert raw-from-sensor .xml file to more convenient .csv file  
```{r Nitratax_XMLtoCSV, include=FALSE}
# Function name:
      # Nitratax_XMLtoCSV
# Function description:
      # Convert output format from Nitratax sensor (.xml) into .csv file
# Input variable definitions:
      # filename = .xml file to be converted, character  string
      # sensor = character string of sensor name, will be outputted in .csv file name
      # deployDate = start date of sensor operation, YYYY-MM-DD character string
# Outputs:
      # .csv file titled "KansasR_Nitratax_<sensor>.csv"

Nitratax_XMLtoCSV <- function(filename, sensor, deployDate){
  
  xmlData <- XML::xmlParse(file = filename)
  xmlData <- XML::xmlToDataFrame(xmlData, homogeneous = F, stringsAsFactors = F)
  xmlData <- xmlData[-(1:6),] # taking out text at top of dataframe
  
  for (count in 1:length(xmlData)){
    names(xmlData)[count] <- as.character(xmlData[1,count])
  }
  
  xmlData <- xmlData[-1,] # taking out character rows
  xmlData <- xmlData[,-1] # taking out first blank column
  
  xmlData$`  TIME  ` <- lubridate::as_datetime(xmlData$`  TIME  `, 
                                               tz = "America/Chicago", 
                                               format = "%m/%d/%Y %H:%M:%S")
  xmlData$`  NITRATE CONC.  ` <- as.numeric(xmlData$`  NITRATE CONC.  `)
  xmlData$`  ER  ` <- as.numeric(xmlData$`  ER  `)
  xmlData$`  EM  ` <- as.numeric(xmlData$`  EM  `)
  
  xml_trunkated <- xmlData[xmlData$`  TIME  `>ymd(deployDate),] # subset only data from deploy onwards
  
  xml_trunkated <- data.frame(xml_trunkated$`  TIME  `, 
                              xml_trunkated$`  NITRATE CONC.  `)
  names(xml_trunkated) <- c("Date and time", "Nitrate plus nitrite as mg-N/L")
  CSV_filename <- paste("KansasR_Nitratax_", 
                        sensor, 
                        ".csv", sep = "")
  
  write.table(xml_trunkated, 
              file = CSV_filename,
              sep = ",", row.names = FALSE)
  return(xml_trunkated)
  
}
```

Import of Nitratax data into R environment, cleanup
```{r NitrataxLoadingCleanup, include=FALSE}
# Variable descriptions:    
      # loadfromCSV = logical vector, if TRUE data will be loaded from .csv files
                                    # if FALSE data will be converted from .xml to .csv
                                    # (computationally much faster to load from .csv)
      # bowersock_N = Nitratax sensor located at the bowersock dam upstream of waste release
      # eric_N = Nitratax sensor located at the waste release point
      # steve_N = Nitratax sensor located about ~5 miles downstream of bowersock

loadfromCSV <- TRUE

if (loadfromCSV == TRUE){
  bowersock_N <- read.csv(file = "KansasR_Nitratax_Bowersock.csv", 
                          header = TRUE,
                          colClasses = c("POSIXct", "numeric"))
  eric_N <- read.csv(file = "KansasR_Nitratax_Eric.csv", 
                     header = TRUE,
                     colClasses = c("POSIXct", "numeric"))
  steve_N <- read.csv(file = "KansasR_Nitratax_Steve.csv", 
                      header = TRUE,
                      colClasses = c("POSIXct", "numeric"))
  
  names(bowersock_N) <- c("Date and time", "Nitrate plus nitrite as mg-N/L")
  names(eric_N) <- c("Date and time", "Nitrate plus nitrite as mg-N/L")
  names(steve_N) <- c("Date and time", "Nitrate plus nitrite as mg-N/L")
  
  cat("data loaded from .csv file")
}
if (loadfromCSV == FALSE) {
  bowersock_N <- Nitratax_XMLtoCSV(filename = "NITRATAX_Bowersock_Pulled05-07-2018.xml", 
                                   sensor = "Bowersock", 
                                   deployDate = "2018-01-13")
  eric_N <- Nitratax_XMLtoCSV(filename = "NITRATAX_Eric_Pulled05-07-2018.xml", 
                              sensor = "Eric", 
                              deployDate = "2018-02-19")
  steve_N <- Nitratax_XMLtoCSV(filename = "NITRATAX_Steve_Pulled05-07-2018.xml", 
                               sensor = "Steve", 
                               deployDate = "2018-01-31")
  
  cat("data loaded from .xml file")
}
# bowersock N cleanup
bowersock_N$`Nitrate plus nitrite as mg-N/L`[bowersock_N$`Date and time` 
                                             %between% 
                                               c(ymd_hms("2018-02-05 21:30:00", tz = "America/Chicago"),
                                                 ymd_hms("2018-02-07 15:30:00", tz = "America/Chicago"))] <- NA

bowersock_N$`Nitrate plus nitrite as mg-N/L`[bowersock_N$`Date and time` 
                                             %between% 
                                               c(ymd_hms("2018-02-16 11:00:00", tz = "America/Chicago"),
                                                 ymd_hms("2018-02-16 11:30:00", tz = "America/Chicago"))] <- NA
# for purposes of sfs talk.....
bowersock_N$`Nitrate plus nitrite as mg-N/L`[bowersock_N$`Date and time` 
                                             %between% 
                                               c(ymd_hms("2018-05-01 00:00:00", tz = "America/Chicago"),
                                                 ymd_hms("2018-05-07 00:00:00", tz = "America/Chicago"))] <- NA

# eric N cleanup
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-02-27 00:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-15 19:30:00", tz = "America/Chicago"))] <- NA
# large precip events
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-18 10:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-18 13:15:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-19 04:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-19 06:00:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-19 08:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-19 10:15:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-19 11:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-19 18:45:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-26 07:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-26 10:30:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-26 11:45:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-26 19:45:00", tz = "America/Chicago"))] <- NA
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-04-26 11:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-04-26 13:00:00", tz = "America/Chicago"))] <- NA
# for purposes of sfs talk.....
eric_N$`Nitrate plus nitrite as mg-N/L`[eric_N$`Date and time` 
                                        %between% 
                                          c(ymd_hms("2018-05-01 00:00:00", tz = "America/Chicago"),
                                            ymd_hms("2018-05-07 00:00:00", tz = "America/Chicago"))] <- NA

# steve N cleanup
steve_N$`Date and time` <- ymd_hms(steve_N$`Date and time`, tz = "America/Chicago")
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-02-13 13:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-02-13 18:45:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-02-14 11:45:00", tz = "America/Chicago"),
                                             ymd_hms("2018-02-14 14:45:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-02-22 12:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-02-27 20:45:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-01 14:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-01 14:30:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-03-09 00:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-03-27 11:30:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-04-12 09:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-04-12 09:30:00", tz = "America/Chicago"))] <- NA
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-04-20 15:15:00", tz = "America/Chicago"),
                                             ymd_hms("2018-04-20 15:45:00", tz = "America/Chicago"))] <- NA
# for purposes of sfs talk.....
steve_N$`Nitrate plus nitrite as mg-N/L`[steve_N$`Date and time` 
                                         %between% 
                                           c(ymd_hms("2018-05-01 00:00:00", tz = "America/Chicago"),
                                             ymd_hms("2018-05-07 00:00:00", tz = "America/Chicago"))] <- NA
```

***  

## miniDOT Data  

Loading miniDOT sensor files: dissolved oxygen and temperature data  
```{r DOloadingCleanup, include=FALSE}
steve_DO <- read.table(file = "miniDOT_Steve_Pulled04-20-2018.txt", header = TRUE,
                       sep = ",", skip = 8, stringsAsFactors = FALSE,
                       colClasses = c("numeric", "POSIXct", "POSIXct", "numeric",
                                      "numeric", "numeric", "numeric", "numeric"),
                       col.names = c("unixTimestamp", "time_UTC", "time_CST", "battery_V",
                                     "temp_C", "DO_mgL", "DO_sat", "Q"))

eric_DO <- read.table(file = "miniDOT_Eric_Pulled05-07-2018.txt", header = TRUE,
                       sep = ",", skip = 8, stringsAsFactors = FALSE,
                       colClasses = c("numeric", "POSIXct", "POSIXct", "numeric",
                                      "numeric", "numeric", "numeric", "numeric"),
                       col.names = c("unixTimestamp", "time_UTC", "time_CST", "battery_V",
                                     "temp_C", "DO_mgL", "DO_sat", "Q"))

# Eric DO cleanup
# 19 - 25 Feb
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-02-19 10:38:00", tz = "America/Chicago"),
                   ymd_hms("2018-02-25 01:28:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-02-19 10:38:00", tz = "America/Chicago"),
                   ymd_hms("2018-02-25 01:28:00", tz = "America/Chicago"))] <- NA
# 5 - 15 Mar
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-05 09:58:00", tz = "America/Chicago"),
                   ymd_hms("2018-03-15 11:08:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-05 09:58:00", tz = "America/Chicago"),
                   ymd_hms("2018-03-15 11:08:00", tz = "America/Chicago"))] <- NA
# 28 Mar
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-28 04:25:00", tz = "America/Chicago"),
                   ymd_hms("2018-03-28 21:25:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-28 04:25:00", tz = "America/Chicago"),
                   ymd_hms("2018-03-28 21:25:00", tz = "America/Chicago"))] <- NA
# 30 Mar - 5 Apr
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-30 23:35:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-05 12:00:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-03-30 23:35:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-05 12:00:00", tz = "America/Chicago"))] <- NA
# 14-18 Apr
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-04-14 12:23:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-18 00:00:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-04-14 12:23:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-18 00:00:00", tz = "America/Chicago"))] <- NA
# 25-30 Apr
eric_DO$DO_mgL[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-04-25 19:33:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-30 15:53:00", tz = "America/Chicago"))] <- NA
eric_DO$temp_C[eric_DO$time_CST 
               %between% 
                 c(ymd_hms("2018-04-25 19:33:00", tz = "America/Chicago"),
                   ymd_hms("2018-04-30 15:53:00", tz = "America/Chicago"))] <- NA

# Steve
# 30 Jan
steve_DO$DO_mgL[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-01-30 15:24:00", tz = "America/Chicago"),
                  ymd_hms("2018-01-30 16:54:00", tz = "America/Chicago"))] <- NA
steve_DO$temp_C[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-01-30 15:24:00", tz = "America/Chicago"),
                  ymd_hms("2018-01-30 16:54:00", tz = "America/Chicago"))] <- NA
# 9-12 Mar
steve_DO$DO_mgL[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-03-09 17:17:00", tz = "America/Chicago"),
                  ymd_hms("2018-03-12 10:47:00", tz = "America/Chicago"))] <- NA
steve_DO$temp_C[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-03-09 17:17:00", tz = "America/Chicago"),
                  ymd_hms("2018-03-12 10:47:00", tz = "America/Chicago"))] <- NA
# 13 Mar
steve_DO$DO_mgL[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-03-13 05:09:00", tz = "America/Chicago"),
                  ymd_hms("2018-03-13 08:49:00", tz = "America/Chicago"))] <- NA
steve_DO$temp_C[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-03-13 05:09:00", tz = "America/Chicago"),
                  ymd_hms("2018-03-13 08:49:00", tz = "America/Chicago"))] <- NA
# 19-20 Mar
steve_DO$DO_mgL[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-04-19 23:37:00", tz = "America/Chicago"),
                  ymd_hms("2018-04-20 08:17:00", tz = "America/Chicago"))] <- NA
steve_DO$temp_C[steve_DO$time_CST 
                %between% 
                c(ymd_hms("2018-04-19 23:37:00", tz = "America/Chicago"),
                  ymd_hms("2018-04-20 08:17:00", tz = "America/Chicago"))] <- NA
```

***

## USGS Data  

Pulling data from USGS server for sensor at Desoto KS (fourth core site)  
```{r DesotoPullCleanup, include=FALSE, warning=FALSE}
siteNoDesoto <- "06892350" # USGS code for Kansas River site at Desoto, KS
pCode <- c("99133", "00060", "00065", "00300", "00301","00010") 
  # USGS parameter codes:
    # 99133 = nitrate + nitrite [mg-N/L]
    # 00060 = discharge [cfs]
    # 00065 = gage height [ft]
    # 00300 = dissolved oxygen [mg/L]
    # 00301 = dissolved oxygen [% saturation]
    # 00010 = water temperature [C]
pullBegin <- as.POSIXct("2018-02-01", format = "%Y-%m-%d") # Date for start of data pull
pullEnd <- as.POSIXct("2018-04-30", format = "%Y-%m-%d") # Date for end of data pull

# USGS data pull
desoto <- readNWISuv(siteNumbers = siteNoDesoto, parameterCd = pCode, 
                     startDate = pullBegin, endDate = pullEnd, tz = "America/Chicago") 

# Rename columns from codes to parameter names
desoto <- renameNWISColumns(desoto)
# Fix remaining names that are still parameter codes
names(desoto)[names(desoto)=="00301_Inst"] <- "DO_percentsat" 
names(desoto)[names(desoto)=="00301_Inst_cd"] <- "DO_percentsat_cd"
names(desoto)[names(desoto)=="99133_Inst"] <- "nitrateNitrite"
names(desoto)[names(desoto)=="99133_Inst_cd"] <- "nitrateNitrite_cd"

# Convert from American to SI units
# cubic foot per second to liters per second
ft3s_m3s <- function(ft3s){
  m3s <- ft3s*0.0283168
  return(m3s)
}
# foot to meters
ft_m <- function(ft){
  m <- 0.3048*ft
  return(m)
}
desoto$Flow_Inst <- ft3s_m3s(desoto$Flow_Inst)
desoto$GH_Inst <- ft_m(desoto$GH_Inst)

# large precipitation events
desoto$nitrateNitrite[desoto$dateTime
                      %between%
                        c(ymd_hms("2018-03-19 20:30:00", tz = "America/Chicago"),
                          ymd_hms("2018-03-21 16:45:00", tz = "America/Chicago"))] <- NA
desoto$nitrateNitrite[desoto$dateTime
                      %between%
                        c(ymd_hms("2018-03-25 00:00:00", tz = "America/Chicago"),
                          ymd_hms("2018-03-30 11:30:00", tz = "America/Chicago"))] <- NA
desoto$nitrateNitrite[desoto$dateTime
                      %between%
                        c(ymd_hms("2018-04-14 13:30:00", tz = "America/Chicago"),
                          ymd_hms("2018-04-14 20:45:00", tz = "America/Chicago"))] <- NA

```

Pulling data from USGS stream gauge at Lawrence, for finding water balance
```{r LawrencePullCleanup, include=FALSE, warning=FALSE}
siteNoLawrence <- "06891080"
pCodesLawrence <- c("00060","00065")

lawrence <- readNWISuv(siteNumbers = siteNoLawrence, parameterCd = pCodesLawrence, 
                       startDate = pullBegin, endDate = pullEnd, tz = "America/Chicago") 

# Rename columns from codes to parameter names
lawrence <- renameNWISColumns(lawrence)

# convert from american to SI units
lawrence$Flow_Inst <- ft3s_m3s(lawrence$Flow_Inst)

lawrence$.Downstream.of.Bowersock.Dam._GH_Inst <- ft_m(lawrence$.Downstream.of.Bowersock.Dam._GH_Inst)

lawrence$.Upstream.of.Bowersock.Dam._GH_Inst <- ft_m(lawrence$.Upstream.of.Bowersock.Dam._GH_Inst)
```

***  

## City of Lawrence Pumping Logs  

Data code         |   Meaning
----------------- |----------
NA                |no sample was taken on this date
BDL_NitrateNitrite|sample concentration was less than 0.25 mg-N/L
BDL_Ammonia       |sample concentration was less than 0.30 mg-N/L

```{r FarmlandLogs_plotting, include=TRUE}
FarmlandPumpingLogs <- read.csv(file = "FarmlandDailyReport_MCKformatted.csv", 
                                header = TRUE, sep = ",")

# conversion from american to metric units
# million gallons to liters
MG_L <- function(MG){
  L <- MG*3.78541*10^6
  return(L)
}

FarmlandPumpingLogs$VolumePumped_MG <- MG_L(FarmlandPumpingLogs$VolumePumped_MG)
names(FarmlandPumpingLogs)[names(FarmlandPumpingLogs) == "VolumePumped_MG"] <- "VolumePumped_L"

# gallons per minute to liters per second
gpm_m3s <- function(gpm){
  m3s <- gpm*6.309E-5
  return(m3s)
}

FarmlandPumpingLogs$flowRate_Outfall001A_gpm <- gpm_m3s(FarmlandPumpingLogs$flowRate_Outfall001A_gpm)
names(FarmlandPumpingLogs)[names(FarmlandPumpingLogs) == "flowRate_Outfall001A_gpm"] <-
  "flowRate_Outfall001A_m3.s"

FarmlandPumpingLogs$flowRate_AveragePump_gpm <- gpm_m3s(FarmlandPumpingLogs$flowRate_AveragePump_gpm)
names(FarmlandPumpingLogs)[names(FarmlandPumpingLogs) == "flowRate_AveragePump_gpm"] <- 
  "flowRate_AveragePump_m3.s"

# Reformatting column data types
FarmlandPumpingLogs$Date <- mdy(FarmlandPumpingLogs$Date)
FarmlandPumpingLogs$Notes <- as.character(FarmlandPumpingLogs$Notes)

LogDateStart <- min(FarmlandPumpingLogs$Date) 
LogDateEnd <- max(FarmlandPumpingLogs$Date)

pullBegin1 <- ymd(pullBegin)
pullEnd1 <- ymd(pullEnd)

FarmlandPump_DrainageDitch_NitratePlot <- ggplot(data = FarmlandPumpingLogs, 
                                                 aes(Date, Downstream_1625RdDitch_NitrateNitrite_mgN.L)) + 
    geom_line(show.legend = FALSE, size = 1, 
              color = wes_palette(n=5, name="Royal2")[5]) +
    coord_cartesian(xlim = c(LogDateStart, LogDateEnd), ylim = c(0, 700)) +
    labs(y = "Nitrate plus nitrite [mg-N/L]", x = "Date", 
         title = "Farmland pipe outflow") +
    theme_classic() + scale_color_manual(values=wes_palette(n=5, name="Darjeeling")[1])

FarmlandPump_DrainageDitch_NitratePlot

FarmlandPump_DrainageDitch_NitratePlot_zoom <- ggplot(data = FarmlandPumpingLogs, 
                                                 aes(Date, Downstream_1625RdDitch_NitrateNitrite_mgN.L)) + 
    geom_point(show.legend = FALSE, size = 1, 
              color = wes_palette(n=5, name="Royal2")[5]) +
    coord_cartesian(xlim = c(pullBegin1, pullEnd1), ylim = c(0, 700)) +
    labs(y = "Nitrate plus nitrite [mg-N/L]", x = "Date", 
         title = "Farmland pipe outflow") +
    theme_classic() + scale_color_manual(values=wes_palette(n=5, name="Darjeeling")[1])

FarmlandPump_DrainageDitch_NitratePlot_zoom
```

***  

## NEON Data   

```{r NEONpullCleanup, include=FALSE, eval=TRUE}
downloadNEONfiles <- FALSE

if (downloadNEONfiles == TRUE){
  dir.create(paste0(getwd(), "/NEONfiles"))
  
  ######## PAR ############
  getPackage(dpID = "DP1.00024.001", site_code = "UKFS",
           year_month = "2018-02", package = "basic",
           savepath = paste(getwd(), "NEONfiles", sep = "/")) # Feb at KU field station
  getPackage(dpID = "DP1.00024.001", site_code = "UKFS",
           year_month = "2018-03", package = "basic",
           savepath = paste(getwd(), "NEONfiles", sep = "/")) # March at KU field station
  #getPackage(dpID = "DP1.00024.001", site_code = "UKFS",
           #year_month = "2018-04", package = "basic",
           #savepath = getwd()) # April at KU field station 
  stackByTable(dpID = "DP1.00024.001", 
               filepath = paste0(getwd(), "/NEONfiles"), 
               savepath = paste0(getwd(), "/NEONfiles"), 
               folder = TRUE)
  
  ######## Barometric pressure ############
  getPackage(dpID = "DP1.00004.001", site_code = "UKFS",
           year_month = "2018-02", package = "basic",
           savepath = paste(getwd(), "NEONfiles", sep = "/")) # Feb at KU field station
  getPackage(dpID = "DP1.00004.001", site_code = "UKFS",
           year_month = "2018-03", package = "basic",
           savepath = paste(getwd(), "NEONfiles", sep = "/")) # March
  #getPackage(dpID = "DP1.00004.001", site_code = "UKFS",
           #year_month = "2018-04", package = "basic",
           #savepath = paste(getwd(), "NEONfiles", sep = "/")) # April
  stackByTable(dpID = "DP1.00004.001", 
               filepath = paste0(getwd(), "/NEONfiles"), 
               savepath = paste0(getwd(), "/NEONfiles"), 
               folder = TRUE)
}

PAR <- read.csv(file = paste0(getwd(), "/NEONfiles", "/stackedFiles", 
                              "/PARPAR_1min.csv"), header = TRUE)
PAR <- subset(PAR, subset = PAR$verticalPosition == 10) # take measurements that are closest to ground
PAR$startDateTime <- ymd_hms(PAR$startDateTime, tz = "America/Chicago")
PAR <- data.frame(
    dateTime = PAR$startDateTime,
    mean = PAR$PARMean,
    min = PAR$PARMinimum,
    max = PAR$PARMaximum,
    var = PAR$PARVariance
)

# PAR units are micromoles per m^2 per second

#### Barometric pressure, units are kPa
BarPress <- read.csv(file = paste0(getwd(), "/NEONfiles", "/stackedFiles", 
                              "/BP_30min.csv"), header = TRUE)
BarPress$startDateTime <- ymd_hms(BarPress$startDateTime, tz = "America/Chicago")
BarPress <- data.frame(
  dateTime = BarPress$startDateTime,
  mean = BarPress$staPresMean,
  min = BarPress$staPresMinimum,
  max = BarPress$staPresMaximum,
  var = BarPress$staPresVariance
)
# convert to mmHg
kpa_mmhg <- function(kpa){
  mmhg <- kpa*7.5006156130264
  return(mmhg)
}
BarPress$mean <- kpa_mmhg(BarPress$mean)
BarPress$min <- kpa_mmhg(BarPress$min)
BarPress$max <- kpa_mmhg(BarPress$max)
BarPress$var <- kpa_mmhg(BarPress$var)
```

***  

# Raw data visualizations  

```{r NmergedPlot, echo=FALSE, warning=FALSE, fig.height=6, fig.width=10}
#combined plot of all N data on one axis
nitrataxMergedPlot <- ggplot(NULL) +
  geom_point(data = bowersock_N, aes(x = `Date and time`, 
                                   y = `Nitrate plus nitrite as mg-N/L`, 
                                   color = "Upstream"), show.legend = TRUE, size = 1) +
  geom_point(data = steve_N, aes(x = `Date and time`, 
                               y = `Nitrate plus nitrite as mg-N/L`, 
                               color = "Downstream"), show.legend = TRUE, size = 1) +
  geom_point(data = eric_N, aes(x = `Date and time`, 
                              y = `Nitrate plus nitrite as mg-N/L`, 
                              color = "Discharge point"), show.legend = TRUE, size = 1) +
  geom_point(data = desoto, aes(x = dateTime, 
                                y = nitrateNitrite, color = "Far downstream"), 
                            show.legend = TRUE, size = 1) +
  coord_cartesian(xlim = c(pullBegin, pullEnd), ylim = c(0, 15), expand = TRUE) +
  labs(y = "Nitrate plus nitrite [mg-N/L]", x = "Date", title = "", color = "Legend") +
  theme_classic() +
  theme(legend.position = "right", legend.direction = "vertical") +
  scale_color_manual(values=wes_palette(n=4, name="Darjeeling"))

nitrataxMergedPlot
```

## DO, Temp, N Plots  

Function setup for individual, simple plots for each site  
```{r simplePlotFunction, include=FALSE}
# Function name:
      # simplePlot
# Function description:
      # function to output a simple plot from specified input variables
# Input variable definitions:
      # dataset = the time series dataset from which the plot will be generated, dataframe
      # xAxis = variable to be plotted on the x axis, POSIXct format
      # yAxis = variable to be plotted on the y axis, numeric
      # plotBegin = date that plot should start, POSIXct format
      # plotEnd = date that plot should end, POSIXct format
      # yMin = minimum value of y axis, numeric
      # yMax = maximum value of y axis, numeric
      # yLabel = desired label of y axis, character string
      # xLabel = desired label of x axis, character string
      # title = desired title of plot, character string
# Outputs:
      # time series plot

simplePlot <- function(dataset, xAxis, yAxis, plotBegin, plotEnd, yMin, yMax, xLabel, yLabel, title){
  ggplot(NULL) + 
    geom_point(data = dataset, aes(x = xAxis, y = yAxis, color = title), 
               show.legend = FALSE, size = 1) +
    coord_cartesian(xlim = c(plotBegin, plotEnd), 
                    ylim = c(yMin, yMax)) +
    labs(y = yLabel, x = xLabel, title = title) +
    theme_classic()
}
```



```{r DO_N_temp_Plots, echo=FALSE, warning=FALSE}
colorBlindPalette <- c("blue2", "red", "darkorchid1", "black")

# bowersock
bowersockPlot_N <- simplePlot(dataset = bowersock_N, 
                              xAxis = bowersock_N$`Date and time`, 
                              yAxis = bowersock_N$`Nitrate plus nitrite as mg-N/L`,
                              plotBegin = pullBegin, plotEnd = pullEnd, xLabel = "Date",
                              yMin = 0, yMax = 1.5, 
                              yLabel = "Nitrate plus nitrite [mg-N/L]",
                              title = "Upstream") +
  scale_color_manual(values=colorBlindPalette[1])

bowersockPlot_N

# Eric
ericPlot_N <- simplePlot(dataset = eric_N, 
                         xAxis = eric_N$`Date and time`,
                         yAxis = eric_N$`Nitrate plus nitrite as mg-N/L`,
                         plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                         yMin = 0, yMax = 20, yLabel = "Nitrate plus nitrite [mg-N/L]",
                         title = "Discharge point") +
  scale_color_manual(values=colorBlindPalette[2])

ericPlot_DO <- simplePlot(dataset = eric_DO, 
                         xAxis = eric_DO$time_CST,
                         yAxis = eric_DO$DO_mgL,
                         plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                         yMin = 0, yMax = 20, yLabel = "Dissolved oxygen [mg/L]",
                         title = "Discharge point") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[2])

ericPlot_temp <- simplePlot(dataset = eric_DO, 
                         xAxis = eric_DO$time_CST,
                         yAxis = eric_DO$temp_C,
                         plotBegin = pullBegin, plotEnd = pullEnd, xLabel = "Date",
                         yMin = -5, yMax = 25, yLabel = "Temperature [C]",
                         title = "Discharge point") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[3])

grid.arrange(ericPlot_N, ericPlot_DO, ericPlot_temp, ncol = 1, nrow = 3)

# Steve (downstream)
stevePlot_N <- simplePlot(dataset = steve_N, 
                          xAxis = steve_N$`Date and time`,
                          yAxis = steve_N$`Nitrate plus nitrite as mg-N/L`,
                          plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                          yMin = 0, yMax = 1.5, yLabel = "Nitrate plus nitrite [mg-N/L]",
                          title = "Downstream (5 km)") + 
  scale_color_manual(values=colorBlindPalette[3])

stevePlot_DO <- simplePlot(dataset = steve_DO, 
                         xAxis = steve_DO$time_CST,
                         yAxis = steve_DO$DO_mgL,
                         plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                         yMin = 0, yMax = 20, yLabel = "Dissolved oxygen [mg/L]",
                         title = "Downstream (5 km)") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[2])

stevePlot_temp <- ericPlot_temp <- simplePlot(dataset = steve_DO, 
                         xAxis = steve_DO$time_CST,
                         yAxis = steve_DO$temp_C,
                         plotBegin = pullBegin, plotEnd = pullEnd, xLabel = "Date",
                         yMin = -5, yMax = 25, yLabel = "Temperature [C]",
                         title = "Downstream (5 km)") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[3])

grid.arrange(stevePlot_N, stevePlot_DO, stevePlot_temp, ncol = 1, nrow = 3)

# Desoto
desotoPlot_N <- simplePlot(dataset = desoto, 
                           xAxis = desoto$dateTime,
                           yAxis = desoto$nitrateNitrite,
                           plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                           yMin = 0, yMax = 1.5, yLabel = "Nitrate plus nitrite [mg-N/L]",
                           title = "Far downstream (30 km)") +
  scale_color_manual(values=colorBlindPalette[4])

desotoPlot_DO <- simplePlot(dataset = desoto, 
                           xAxis = desoto$dateTime,
                           yAxis = desoto$DO_Inst,
                           plotBegin = pullBegin, plotEnd = pullEnd, xLabel = NULL,
                           yMin = 0, yMax = 20, yLabel = "Dissolved oxygen [mg/L]",
                           title = "Far downstream (30 km)") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[2])

desotoPlot_temp <- simplePlot(dataset = desoto, 
                              xAxis = desoto$dateTime,
                              yAxis = desoto$Wtemp_Inst,
                              plotBegin = pullBegin, plotEnd = pullEnd, xLabel = "Date",
                              yMin = -5, yMax = 25, yLabel = "Temperature [C]",
                              title = "Far downstream (30 km)") +
  theme(plot.title = element_blank()) +
  scale_color_manual(values=wes_palette(n=3, name="Darjeeling")[3])

grid.arrange(desotoPlot_N, desotoPlot_DO, desotoPlot_temp, ncol = 1, nrow = 3)
```

## N data plot (include in talk)  

```{r StackedNPlot, fig.height=9, fig.width=9}
stackedNPlot <- 
  grid.arrange(
    bowersockPlot_N +
      coord_cartesian(xlim = c(pullBegin, pullEnd),
                      ylim = c(0, 1.5)) +
      xlab(NULL) +
      ylab(NULL) +
      ggtitle(NULL) +
      theme(axis.text = element_text(size = 14),
            axis.text.x = element_blank()) +
      scale_color_manual(values = colorBlindPalette[1]) + 
      scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
      geom_vline(xintercept = ymd_hms("2018-04-01 00:00:00")),
  
    ericPlot_N +
      coord_cartesian(xlim = c(pullBegin, pullEnd),
                      ylim = c(0, 10)) +
      xlab(NULL) +
      ylab(NULL) +
      ggtitle(NULL) +
      theme(axis.text = element_text(size = 14),
            axis.text.x = element_blank()) +
      scale_color_manual(values = colorBlindPalette[2]) +
      scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
      geom_vline(xintercept = ymd_hms("2018-04-01 00:00:00")),
  
    stevePlot_N +
      coord_cartesian(xlim = c(pullBegin, pullEnd),
                      ylim = c(0, 1.5)) +
      xlab(NULL) +
      ylab(NULL) +
      ggtitle(NULL) +
      theme(axis.text = element_text(size = 14),
            axis.text.x = element_blank()) +
      scale_color_manual(values = colorBlindPalette[3]) +
      scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
      geom_vline(xintercept = ymd_hms("2018-04-01 00:00:00")),
  
    desotoPlot_N +
      coord_cartesian(xlim = c(pullBegin, pullEnd),
                      ylim = c(0, 1.5)) +
      xlab(NULL) +
      ylab(NULL) +
      ggtitle(NULL) +
      theme(axis.text = element_text(size = 14)) +
      scale_color_manual(values = colorBlindPalette[4]) +
      scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
      geom_vline(xintercept = ymd_hms("2018-04-01 00:00:00")),
  
    nrow = 4,
    ncol = 1
)

ggsave(filename = "StackedNPlot.png", plot = stackedNPlot, dpi = 500)
```

## Combined temperature plot (include in talk)
```{r comboTemp, fig.width=9, fig.height=3}
comboTemp <- 
ggplot() +
  geom_point(data = desoto, aes(x = dateTime, y = Wtemp_Inst, color = "site 4"),
             show.legend = FALSE) +
  geom_point(data = steve_DO, aes(x = time_CST, y = temp_C, color = "site 2"),
             show.legend = FALSE) +
  geom_point(data = eric_DO, aes(x = time_CST, y = temp_C, color = "site 3"),
             show.legend = FALSE) +
  coord_cartesian(xlim = c(pullBegin, pullEnd), 
                  ylim = c(-5, 25)) +
  theme_classic() +
  scale_color_manual(values = colorBlindPalette[2:4]) +
  ylab("Water temperature, deg C") +
  xlab(NULL) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14)) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00"))

ggsave(filename = "temp_comboplot.png", plot = comboTemp, dpi = 500,
       width = 9, height = 3, units = "in")
```


## PAR plot (include in talk)
```{r PARplotting, fig.height=3, fig.width=9, eval=TRUE}
# calculate sum of PAR mean values
PAR$PARMean_sum <- cumsum(PAR$mean)

# convert PAR units (from umol m^-2 s^-1 to mol m^-2 d^-1)
umolm2s_to_molm2d <- function(umol_m2_s){
  mol_m2_s <- umol_m2_s*1.0E-6
  mol_m2_d <- mol_m2_s*86400
  return(mol_m2_d)
}

PAR$PARMean_sum_molm2d <- umolm2s_to_molm2d(PAR$PARMean_sum)

#convert PAR units (from umol m^-2 s^-1 to mol m^-2 s^-1)
umol_mol <- function(umol){
  mol <- umol*1.0E-6
  return(mol)
}

PAR$PARMean_sum_molm2s <- umol_mol(PAR$PARMean_sum)

#ok our raw values are on the same magnitide as the Lupon paper, but the converted values are not similar at all, wth

PAR_rawPlot <- 
  ggplot(NULL) + 
    geom_line(data = PAR, aes(x = dateTime, y = mean), 
               show.legend = FALSE, size = 1) +
    coord_cartesian(xlim = c(pullBegin, pullEnd)) +
    labs(y = "PAR, umol m^-2 s^-1", 
         x = NULL) +
    theme_classic() +
    theme(axis.text = element_text(size = 14),
          axis.title = element_text(size = 14)) +
    scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
    geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00"))

PAR_sumPlot <- 
  ggplot(NULL) + 
    geom_area(data = PAR, aes(x = startDateTime, y = PARMean_sum), 
               show.legend = FALSE, size = 1) +
    coord_cartesian(xlim = c(pullBegin, pullEnd)) +
    labs(y = "Sum PAR, 
         umol m^-2 s^-1", 
         x = "Date") +
    theme_classic() +
    theme(axis.text = element_text(size = 14),
          axis.title = element_text(size = 14))

PAR_sumPlot_convertunits_molday <- 
  ggplot(NULL) + 
    geom_area(data = PAR, aes(x = startDateTime, y = PARMean_sum_molm2d), 
               show.legend = FALSE, size = 1) +
    coord_cartesian(xlim = c(pullBegin, pullEnd)) +
    labs(y = "Sum PAR, 
         mol m^-2 d^-1", 
         x = "Date") +
    theme_classic() +
    theme(axis.text = element_text(size = 14),
          axis.title = element_text(size = 14))

PAR_sumPlot_convertunits_mol <- 
  ggplot(NULL) + 
    geom_area(data = PAR, aes(x = startDateTime, y = PARMean_sum_molm2s), 
               show.legend = FALSE, size = 1) +
    coord_cartesian(xlim = c(pullBegin, pullEnd)) +
    labs(y = "Sum PAR, 
         mol m^-2 s^-1", 
         x = "Date") +
    theme_classic() +
    theme(axis.text = element_text(size = 14),
          axis.title = element_text(size = 14))

PAR_rawPlot

ggsave(filename = "PAR_rawPlot.png", plot = PAR_rawPlot, dpi = 500,
       width = 9, height = 3, units = "in")
```

## Time period plots  

```{r PeriodPlots, include=TRUE, fig.height=5, fig.width=6}
# Bowersock zoom N plot
bowersockPlot_N +
  coord_cartesian(xlim = c(mdy_hms("02-21-2018 12:00:00"),mdy_hms("02-27-2018 12:00:00")),
                  ylim = c(0,0.4)) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  ylab(NULL) +
  xlab(NULL) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))

# eric zoom N plot
ericPlot_N +
    coord_cartesian(xlim = c(mdy_hms("03-20-2018 12:00:00"),mdy_hms("03-26-2018 12:00:00")),
                    ylim = c(0,10)) +
    scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
    ylab(NULL) +
    xlab(NULL) +
    ggtitle(NULL) +
    theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))

# steve zoom N plot
stevePlot_N +
    coord_cartesian(xlim = c(mdy_hms("02-10-2018 12:00:00"),mdy_hms("02-16-2018 12:00:00")),
                    ylim = c(0,1)) +
    scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
    ylab(NULL) +
    xlab(NULL) +
    ggtitle(NULL) +
    theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))

# desoto zoom N plot
desotoPlot_N +
    coord_cartesian(xlim = c(mdy_hms("03-10-2018 12:00:00"),mdy_hms("03-16-2018 12:00:00")),
                    ylim = c(0,1)) +
    scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
    ylab(NULL) +
    xlab(NULL) +
    ggtitle(NULL) +
    theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))
```

# Water Balance  

```{r waterBalance, fig.height=3, fig.width=10, include=TRUE}
#subset data frames
desoto_sub <- data.frame(desoto$dateTime, desoto$Flow_Inst)
names(desoto_sub) <- c("dateTime", "Flow_Inst")
desoto_sub$dateTime <- ymd_hms(desoto_sub$dateTime)

lawrence_sub <- data.frame(lawrence$dateTime, lawrence$Flow_Inst)
names(lawrence_sub) <- c("dateTime", "Flow_Inst")
lawrence_sub$dateTime <- ymd_hms(lawrence_sub$dateTime)

# combine data frames
waterBal <- merge(desoto_sub, lawrence_sub, 
                  by = "dateTime", suffixes = c(".desoto", ".lawrence"))

#calculate difference between downstream and upstream gauge (balance = downstream discharge - upstream)
waterBal$difference <- waterBal$Flow_Inst.desoto - waterBal$Flow_Inst.lawrence

# plotting flow at upstream and downstream locations on same time series graph
ggplot() +
  geom_line(data = desoto, aes(x = dateTime, y = Flow_Inst, color = "Desoto")) +
  geom_line(data = lawrence, aes(x = dateTime, y = Flow_Inst, color = "Lawrence"))

# plot the water balance graph centered on 0: 
      # positive values = gaining stream
      # negative values = losing stream
waterBal_plot <- ggplot() +
  geom_line(data = waterBal, 
            aes(x = dateTime, y = difference, color = "difference"),
            show.legend = FALSE, 
            color = colorBlindPalette[4], size = 1.5) +
  geom_hline(aes(yintercept = mean(na.omit(waterBal$difference))),
             color = "red",
             linetype = "solid") +
  coord_cartesian(xlim = c(pullBegin, pullEnd), 
                  ylim = c(-50, 50)) +
  labs(y = "Difference between downstream and upstream flow, m3/s",
       x = NULL) +
  theme_classic() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14)) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
    geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00"))

waterBal_plot

ggsave(filename = "waterBal_plot.png", plot = waterBal_plot, dpi = 500,
       width = 9, height = 3, units = "in")
```

***  

## Temporal difference in N (include in talk)  

```{r Ntemporal, fig.height=9, fig.width=9}
# bowersock
bowersock_N_0 <- bowersock_N[grep("00:00:00", bowersock_N$`Date and time`),]
bowersock_N_12 <- bowersock_N[grep("12:00:00", bowersock_N$`Date and time`),]
bowersockN_daynightdiff <- merge(
  bowersock_N_0,
  bowersock_N_12,
  by = c("Date and time", "Nitrate plus nitrite as mg-N/L"), all = TRUE
)
bowersockN_daynightdiff$differences <- c(NA, 
                                         diff(bowersockN_daynightdiff$`Nitrate plus nitrite as mg-N/L`, 
                                         differences = 1))
bowersockN_daynightdiff$perc <- NA
for (i in 2:nrow(bowersockN_daynightdiff)){
    bowersockN_daynightdiff$perc[i] <- 
      bowersockN_daynightdiff$differences[i] / 
      bowersockN_daynightdiff$`Nitrate plus nitrite as mg-N/L`[i-1]
}
bowersockN_daynightdiff$perc[bowersockN_daynightdiff$`Date and time`
                             %between%
                               c(ymd_hms("2018-02-27 05:00:00"), 
                                 ymd_hms("2018-04-30 14:45:00"))] <- 0

# eric
eric_N_0 <- eric_N[grep("00:00:00", eric_N$`Date and time`),]
eric_N_12 <- eric_N[grep("12:00:00", eric_N$`Date and time`),]
ericN_daynightdiff <- merge(
  eric_N_0,
  eric_N_12,
  by = c("Date and time", "Nitrate plus nitrite as mg-N/L"), all = TRUE
)
ericN_daynightdiff$differences <- c(NA, 
                                    diff(ericN_daynightdiff$`Nitrate plus nitrite as mg-N/L`, 
                                    differences = 1))
ericN_daynightdiff$perc <- NA
for (i in 2:nrow(ericN_daynightdiff)){
    ericN_daynightdiff$perc[i] <- 
    ericN_daynightdiff$differences[i] / 
    ericN_daynightdiff$`Nitrate plus nitrite as mg-N/L`[i-1]
}

# steve
steve_N_0 <- steve_N[grep("00:00:00", steve_N$`Date and time`),]
steve_N_12 <- steve_N[grep("12:00:00", steve_N$`Date and time`),]
steveN_daynightdiff <- merge(
  steve_N_0,
  steve_N_12,
  by = c("Date and time", "Nitrate plus nitrite as mg-N/L"), all = TRUE
)
steveN_daynightdiff$differences <- c(NA, 
                                    diff(steveN_daynightdiff$`Nitrate plus nitrite as mg-N/L`, 
                                    differences = 1))
steveN_daynightdiff$perc <- NA
for (i in 2:nrow(steveN_daynightdiff)){
    steveN_daynightdiff$perc[i] <- 
    steveN_daynightdiff$differences[i] / 
    steveN_daynightdiff$`Nitrate plus nitrite as mg-N/L`[i-1]
}

# desoto
desoto_N_0 <- desoto[grep("00:00:00", desoto$dateTime),]
desoto_N_12 <- desoto[grep("12:00:00", desoto$dateTime),]
desotoN_daynightdiff <- merge(
  desoto_N_0,
  desoto_N_12,
  by = c("dateTime", "nitrateNitrite"), all = TRUE
)
desotoN_daynightdiff$differences <- c(NA, 
                                      diff(desotoN_daynightdiff$nitrateNitrite, 
                                      differences = 1))
desotoN_daynightdiff$perc <- NA
for (i in 2:nrow(desotoN_daynightdiff)){
    desotoN_daynightdiff$perc[i] <- 
    desotoN_daynightdiff$differences[i] / 
    desotoN_daynightdiff$nitrateNitrite[i-1]
}

# arrange all, y label is "Change in [NOx] (%)"
ndiff_plot <- 
grid.arrange(
  ggplot(bowersockN_daynightdiff, aes(x = `Date and time`, y = perc, color = "Site 4")) +
  geom_line(size = 2) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,2), x = c(mdy_hms("02/01/2018 00:00:00"), mdy_hms("05/01/2018 00:00:00"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.position = "none", axis.text.x = element_blank()) +
  scale_color_manual(breaks = "Site 4", values = colorBlindPalette[1]) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00")) +
  geom_hline(yintercept = 0),
  
  ggplot(ericN_daynightdiff, aes(x = `Date and time`, y = perc, color = "Site 2")) +
  geom_line(size = 2) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,2), x = c(mdy_hms("02/01/2018 00:00:00"), mdy_hms("05/01/2018 00:00:00"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.position = "none", axis.text.x = element_blank()) +
  scale_color_manual(breaks = "Site 4", values = colorBlindPalette[2]) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00")) +
  geom_hline(yintercept = 0),
  
  ggplot(steveN_daynightdiff, aes(x = `Date and time`, y = perc, color = "Site 4")) +
  geom_line(size = 2) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,2), x = c(mdy_hms("02/01/2018 00:00:00"), mdy_hms("05/01/2018 00:00:00"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.position = "none", axis.text.x = element_blank()) +
  scale_color_manual(breaks = "Site 4", values = colorBlindPalette[3]) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00")) +
  geom_hline(yintercept = 0),
  
  ggplot(desotoN_daynightdiff, aes(x = dateTime, y = perc, color = "Site 4")) +
  geom_line(size = 2) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,2), x = c(mdy_hms("02/01/2018 00:00:00"), mdy_hms("05/01/2018 00:00:00"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.position = "none") +
  scale_color_manual(breaks = "Site 4", values = colorBlindPalette[4]) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00")) +
  geom_hline(yintercept = 0),
  
  nrow = 4,
  ncol = 1
)

ggsave(filename = "Ndiff.png", plot = ndiff_plot, dpi = 500)

```

```{r smaller section of N diff, fig.height=5, fig.width=6}
# smaller section
ggplot(desotoN_daynightdiff, aes(x = dateTime, y = perc, color = "Site 4")) +
  geom_line(size = 2) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,2), x = c(mdy_hms("03-17-2018 00:00:00"),
                                     mdy_hms("03-19-2018 12:00:00"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.position = "none") +
  scale_color_manual(breaks = "Site 4", values = colorBlindPalette[4]) +
  scale_x_datetime(date_breaks = "day", date_labels = "%b %d") +
  geom_vline(xintercept = mdy_hms("04/01/2018 00:00:00")) +
  geom_hline(yintercept = 0)

desotoPlot_N +
    coord_cartesian(xlim = c(mdy_hms("03-17-2018 00:00:00"),
                             mdy_hms("03-19-2018 12:00:00")),
                    ylim = c(0,0.75)) +
    scale_x_datetime(date_breaks = "day", date_labels = "%b %d") +
    ylab(NULL) +
    xlab(NULL) +
    ggtitle(NULL) +
    theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14))
```

# Metabolism modeling

```{r StreamMetabolizer}
# round datetime_solar
#desoto$dateTime_solar <- round_date(desoto$dateTime_solar, "10 minutes")

# estimate stream depth from discharge data
  # using d = c*Q^f Leopold and Maddock 1953
  # default values for c and f as suggested in Raymond et al 2012
      # c = 0.409 m, f = 0.294 **verify if these are representative of Kaw
desoto$depth <- calc_depth(desoto$Flow_Inst)
  # this gives us close the the GH readings for depth

# assemble data frame of data needed for metabolism model
desoto_MetabData <- merge.data.frame(desoto, PAR, by = "dateTime")
# calculate solar time
desoto_longitude <- -94.964616
desoto_MetabData$dateTime_solar <- calc_solar_time(desoto_MetabData$dateTime, 
                                                   longitude = desoto_longitude)
# cleanup data frame 
desoto_MetabData <- 
data.frame(solar.time = desoto_MetabData$dateTime_solar,
           temp.water = desoto_MetabData$Wtemp_Inst,
           discharge = desoto_MetabData$Flow_Inst,
           DO.obs = desoto_MetabData$DO_Inst,
           DO.sat = desoto_MetabData$DO_percentsat,
           depth = desoto_MetabData$depth,
           light = desoto_MetabData$mean)

# model fitting
    # using the basic MLE model
mle_name <- mm_name(type='mle', ode_method = "trapezoid")
mle_specs <- specs(mle_name)
mle_specs <- revise(mle_specs, day_start = 4, day_end = 28, required_timestep = NA,
                    split_dates = TRUE, verbose = TRUE)
mm_mle_desoto <- metab(mle_specs, data = desoto_MetabData)

# model plotting
plot_metab_preds(mm_mle_desoto)
plot_DO_preds(mm_mle_desoto)
# still having gapping between dates

# view errors
get_params(mm_mle_desoto) %>% select(date, warnings, errors)

# trying a bayseian model for better prediction of multi-day metabolism
bayes_name <- mm_name(type = "bayes", pool_K600 = "binned")
bayes_specs <- specs(bayes_name) 
bayes_specs <- revise(bayes_specs, burnin_steps=500, saved_steps=500,
                      split_dates = FALSE, verbose = TRUE, GPP_fun = "limlight",
                      ER_fun = "q10temp", deficit_src = "DO_obs", ode_method = "trapezoid",
                      control = list(max_treedepth = 30))
mm_bayes_desoto <- metab_bayes(specs = bayes_specs, data = desoto_MetabData)
mm_bayes_desoto_stanobj <- get_mcmc(mm_bayes_desoto)

traceplot(mm_bayes_desoto_stanobj)
get_fit(mm_bayes_desoto)$warnings


get_params(mm_bayes_desoto) %>% select(date, warnings, errors)

plot_metab_preds(mm_bayes_desoto)
plot_DO_preds(mm_bayes_desoto)

# what if we estimate K first, then find GPP and ER with fixed K
kmodel_name <- mm_name(type = "Kmodel", engine = "mean")
kmodel_desoto <- metab_Kmodel(specs = kmodel_name, data = desoto_MetabData)
```

```{r MetabolismModel_HotchkissHall_Equations}
# simpler hotchkiss & hall approach, let's start with steve [supp materials from lamberti textbook]

# [2] calculate oxygen saturation at temperature (the minidot already does this)
osat<- function(temp, bp) {
	
	tstd<-log((298.15-temp) / (273.15 + temp))
	
	a0<-2.00907
	a1<-3.22014
	a2<-4.0501
	a3<-4.94457
	a4<- -0.256847
	a5<- 3.88767
	
	u<-10^(8.10765-(1750.286/(235+temp)))
	
	sato<-(exp(a0 + a1*tstd + a2*tstd^2 + a3*tstd^3 + a4*tstd^4+a5*tstd^5))*((bp-u)/(760-u))*1.42905
	sato
}

# [3] calculate barometric pressure

## Function to correct barometric pressure for altitude. From Colt (2012).
## This function gives bp in mmHg for altitude given nearby measurement of standardized barometric pressure. 
	## temp is degC 
	## alt is m
	## bpst is in inches of Hg and the sort of data you will find from U.S. weather stations.  Delete the *25.4 if you in the rest of the world where pressure is given in metric units
####function returns mm of Hg
bpcalc<- function(bpst, alt) {
  bpst*25.4*exp((-9.80665*0.0289644*alt)/(8.31447*(273.15+15)))
}

# [4] load gas exchange (K) functions

## This code estimates K600 for KO2 at a given temperature. From Wanninkhof (1992).
K600fromO2<-function (temp, KO2) {
    ((600/(1800.6 - (120.1 * temp) + (3.7818 * temp^2) - (0.047608 * temp^3)))^-0.5) * KO2
}

# [5] load river metabolism function 

# parameter names (and units)
	# oxy.mod = modeled O2 (mg/L)
	# oxy = O2 data (mg/L)
	# MET[1] = GPP = estimated daily gross primary production (g O2 m-2 d-1); + flux of O2 production
	# MET[2] = ER = estimated daily ecosystem respiration (g O2 m-2 d-1); - flux of O2 consumption 
	# MET[3] = K = estimated K600 (d-1)
	# z = estimated site depth (m)
	# Kcor = KO2 corrected for temperature, calculated from K600 (d-1)
	# bp = barometric pressure (mmHg)
	# temp = water temperature (C)
	# light = PAR data (or modlight from above light function)
	# ts = time step of O2 data from logger (10 min intervals --> units are day, so 10/1440=0.06944)

Kcor<-function (temp,K600) {
	K600/(600/(1800.6-(temp*120.1)+(3.7818*temp^2)-(0.047608*temp^3)))^-0.5
}

# Model to calculate GPP, ER and K simultaneously
  # This model is advantageous in the sense that one can estimate K from the data, 
  # but beware that it may be an overparameterized model.  

# rivermetabK is the master function that calls the MLE function (onestationmleK) and plotting function (onestationplot) below
rivermetabK<-function(o2file, z, bp, ts){
  temp<-o2file$temp
  oxy<-o2file$oxy
  light<-o2file$light

  ##This calls onestationmleK (below) to calculate metabolism by non-linear minimization. We use nlm() function in R stats package; for more information see "help(nlm)"  The first argument is the function to be minimized, the second defines the starting values.  The function that is minimized (onestationmleK, see below) always has as its first argument, a vector of parameters (MET), and second argument a vector of starting values for those parameters, p=c(3,-5, 10).

  river.mle<-nlm(onestationmleK, p=c(3,-5, 10), 
                 oxy=oxy, z=z, temp=temp, 
                 light=light, bp=bp, ts=ts)

  ##plot modeled and measaured O2 given MLE estimates of GPP, ER, and K600.  It calls a function below onestationplot()

  mle_estimates <- onestationplot(GPP = river.mle$estimate[1],
                 ER = river.mle$estimate[2],
                 oxy=oxy, z=z, temp=temp, light=light, 
                 K=river.mle$estimate[3], bp=bp, ts=ts)
  
  ##return GPP, ER, K600, and MLE values
	b <- list(GPP = river.mle$estimate[1],
	          ER = river.mle$estimate[2], 
	          K600 = river.mle$estimate[3], 
	          neglogL = river.mle$minimum[1])
	b
}

# This function returns the negative log likelihood value given O2 data and estimates of GPP, ER, and K (which is vector MET); is included in master rivermetabK function above
onestationmleK <- function(MET, temp, oxy, light, z, bp, ts) {

  oxy.mod <- numeric(length(data))
  # give starting value from oxygen data; this is the only time O2 data is used to model GPP and ER
  oxy.mod[1] <- oxy[1]

  # this is the metabolism equation as in Van de Bogert et al 2007 L&OMethods
  for (i in 2:length(oxy)) {
    oxy.mod[i] <- oxy.mod[i-1] + 
                  ((MET[1]/z) * (light[i]/sum(light))) +
                  MET[2]*ts/z + 
                  (Kcor(temp[i],MET[3])) * ts * (osat(temp[i],bp) - oxy.mod[i-1]) 
    }

  # below is MLE calculation; output is -log likelihood
  # diff in measured and modeled O2
  sqdiff <- (oxy-oxy.mod)^2

  # likelihood function
  L <- length(oxy) * (log(((sum(sqdiff)/length(oxy))^0.5)) + 
       0.5*log(6.28)) + 
       ((2*sum(sqdiff)/length(oxy))^-1) * sum(sqdiff)
	L
}

# this function plots modeled O2 and O2 data from estimates of daily GPP, ER, and K; is included in master rivermetabK function above
# Calls same metabolism equation as in mle function, but plots modeled O2 as a function of GPP, ER, and K estimates from mle
# use this to visually assess your model estimates (should be good agreement between modeled and measured O2)
onestationplot <- function(GPP, ER, oxy, z, temp, K, light, bp, ts) {

  oxy.mod<-numeric(length(oxy))
  oxy.mod[1]<-oxy[1]

  # this is the metabolism equation as in Van de Bogert et al (2007) L&OMethods
  for (i in 2:length(oxy)) { 
    oxy.mod[i] <- oxy.mod[i-1] + 
                  ((GPP/z)*(light[i]/sum(light))) +
                  ER*ts/z + 
                  (Kcor(temp[i],K)) * ts * (osat(temp[i],bp)-oxy.mod[i-1]) 
    }

  plot(seq(1:length(oxy)),
       oxy.mod, 
       type="l", xlab="Time", ylab="Dissolved oxygen  (mg/L)", 
       cex.lab=1.5, cex.axis=1.5, lwd=2)
  points(seq(1:length(oxy)), oxy)
}
```

```{r MetabolismModel_HotchkissHall_Evaluation}
# [1] load data, date and time must be in different columns
eric_metab <- 
data.frame(
  dateTime = ymd_hms(eric_DO$time_CST, tz = "America/Chicago")
)
steve_metab <- 
data.frame(
  dateTime = ymd_hms(steve_DO$time_CST, tz = "America/Chicago")
)
desoto_metab <- 
  data.frame(
  dateTime = ymd_hms(desoto$dateTime, tz = "America/Chicago")
)

# [2] calculate oxygen saturation at temperature (the minidot already does this)
eric_metab$DOmgL <- eric_DO$DO_mgL
eric_metab$temp <- eric_DO$temp_C

steve_metab$DOmgL <- steve_DO$DO_mgL
steve_metab$temp <- steve_DO$temp_C

desoto_metab$DOmgL <- desoto$DO_Inst
desoto_metab$temp <- desoto$Wtemp_Inst

# [3] calculate barometric pressure
eric_alt <- 246.89 # m
eric_pressure <- 
  data.frame(
    dateTime = BarPress$dateTime,
    mean_pressure = BarPress$mean
  )
steve_alt <- 249.02 # m
steve_pressure <- 
  data.frame(
    dateTime = BarPress$dateTime,
    mean_pressure = BarPress$mean
  )
desoto_alt <- 229.78 # m
desoto_pressure <- 
  data.frame(
    dateTime = BarPress$dateTime,
    mean_pressure = BarPress$mean
  )

# [4] load gas exchange (K) functions
eric_metab$K600 <- K600fromO2(temp = eric_metab$temp, KO2 = eric_metab$DOmgL)
steve_metab$K600 <- K600fromO2(temp = steve_metab$temp, KO2 = steve_metab$DOmgL)
desoto_metab$K600 <- K600fromO2(temp = desoto_metab$temp, KO2 = desoto_metab$DOmgL)

# [5] depth
#     using streamMetabolizer package and desoto discharge data here. this won't be entirely accurate.
# estimate stream depth from discharge data
  # using d = c*Q^f Leopold and Maddock 1953
  # default values for c and f as suggested in Raymond et al 2012
      # c = 0.409 m, f = 0.294 **verify if these are representative of Kaw
eric_depth <- 
  data.frame(
    dateTime = ymd_hms(desoto$dateTime, tz = "America/Chicago"),
    z = calc_depth(desoto$Flow_Inst)
  )
steve_depth <- 
  data.frame(
    dateTime = ymd_hms(desoto$dateTime, tz = "America/Chicago"),
    z = calc_depth(desoto$Flow_Inst)
  )
desoto_depth <- 
  data.frame(
    dateTime = ymd_hms(desoto$dateTime, tz = "America/Chicago"),
    z = calc_depth(desoto$Flow_Inst)
  )

# [6] combine metabolism data frame into one, while making sure dates match

# need to combine steve_metab with steve_pressure, steve_depth, and PAR
steve_metab$dateTime <- round_date(steve_metab$dateTime, "10 minutes") # round to nearest 10 min interval
eric_metab$dateTime <- round_date(eric_metab$dateTime, "10 minutes") 

# merge function
merge_metab <- function(metabDataframe, pressureDataframe, depthDataframe, rounddateTime){
  if (rounddateTime == TRUE){
    metabDataframe$dateTime <- round_date(metabDataframe$dateTime, "10 minutes")
  }
  metabDataframe <- 
    merge(metabDataframe,
          pressureDataframe,
          by = "dateTime")
  metabDataframe <- 
    merge(metabDataframe,
          depthDataframe,
          by = "dateTime")
  metabDataframe <- 
    merge(metabDataframe,
          PAR,
          by = "dateTime")
  metabDataframe <- 
    data.frame(dateTime = metabDataframe$dateTime,
               oxy = metabDataframe$DOmgL,
               temp = metabDataframe$temp,
               K600 = metabDataframe$K600,
               bp = metabDataframe$mean_pressure,
               z = metabDataframe$z,
               light = metabDataframe$mean)
  metabDataframe
}

eric_metab <- merge_metab(eric_metab, eric_pressure, eric_depth, TRUE)
steve_metab <- merge_metab(steve_metab, steve_pressure, steve_depth, TRUE)
desoto_metab <- merge_metab(desoto_metab, desoto_pressure, desoto_depth, FALSE)

# [7] run metabolism model
eric_metab_output <- rivermetabK(o2file = na.omit(eric_metab), 
                                  z = 0.2, 
                                  bp = mean(eric_metab$bp, na.rm = TRUE), 
                                  ts = 0.02083333)
eric_metab_output
steve_metab_output <- rivermetabK(o2file = na.omit(steve_metab), 
                                  z = 0.2, 
                                  bp = mean(steve_metab$bp, na.rm = TRUE), 
                                  ts = 0.02083333)
steve_metab_output
desoto_metab_output <- rivermetabK(o2file = na.omit(desoto_metab),
                                   z = 0.2,
                                   bp = mean(desoto_metab$bp, na.rm = TRUE),
                                   ts = 0.02083333)
desoto_metab_output

# function to subset metabolism data and run model for specific time period
metab_timeperiods <- function(metab_dataframe, startTime, endTime){
  metab_timeperiod_subsection <- metab_dataframe[metab_dataframe$dateTime 
                                                 %between% 
                                                   c(ymd_hms(startTime, tz = "America/Chicago"),
                                                     ymd_hms(endTime, tz = "America/Chicago")),]
  metab_output <- rivermetabK(o2file = metab_timeperiod_subsection,
                              z = 0.2,
                              bp = mean(metab_timeperiod_subsection$bp, na.rm = TRUE),
                              ts = 0.02083333)
  return(metab_output)
}

# eric
metabmodels_eric <- 
  list(
    feb_26 = metab_timeperiods(eric_metab, "2018-02-25 02:00:00", "2018-02-27 02:00:00"), # Feb 25-27
    mar_03 = metab_timeperiods(eric_metab, "2018-03-03 09:30:00", "2018-03-05 09:30:00"), # Mar 2-4
    mar_23 = metab_timeperiods(eric_metab, "2018-03-22 00:00:00", "2018-03-24 00:00:00"), # Mar 22-24
    mar_29 = metab_timeperiods(eric_metab, "2018-03-28 22:00:00", "2018-03-30 23:30:00"), # Mar 28-30
    apr_06 = metab_timeperiods(eric_metab, "2018-04-05 12:00:00", "2018-04-07 12:00:00"), # Apr 05-07
    apr_09 = metab_timeperiods(eric_metab, "2018-04-08 00:00:00", "2018-04-09 12:00:00"), # Apr 08-10
    apr_12 = metab_timeperiods(eric_metab, "2018-04-11 17:30:00", "2018-04-13 17:30:00"), # Apr 11-13
    apr_21 = metab_timeperiods(eric_metab, "2018-04-20 00:00:00", "2018-04-22 00:00:00"), # Apr 20-22
    apr_24 = metab_timeperiods(eric_metab, "2018-04-23 19:00:00", "2018-04-25 19:00:00") # Apr 23-25
    )
metabmodels_eric_results <- 
  data.frame(
    GPP = t(as.data.frame(lapply(metabmodels_eric, `[[`, 1))),
    ER = t(as.data.frame(lapply(metabmodels_eric, `[[`, 2)))
  )
metabmodels_eric_results$date <- strptime(row.names(metabmodels_eric_results), format = "%b_%d")
metabmodels_eric_results$date <- as.Date(metabmodels_eric_results$date)

# steve
metabmodels_steve <- 
  list(
    feb_05 = metab_timeperiods(steve_metab, "2018-02-04 00:00:00", "2018-02-06 00:00:00"), # Feb 4-6
    feb_12 = metab_timeperiods(steve_metab, "2018-02-11 00:00:00", "2018-02-13 00:00:00"), # Feb 11-13
    feb_21 = metab_timeperiods(steve_metab, "2018-02-20 00:00:00", "2018-02-22 00:00:00"), # Feb 20-22
    feb_26 = metab_timeperiods(steve_metab, "2018-02-25 02:00:00", "2018-02-27 02:00:00"), # Feb 25-27
    mar_03 = metab_timeperiods(steve_metab, "2018-03-02 09:30:00", "2018-03-04 09:30:00"), # Mar 2-4
    mar_08 = metab_timeperiods(steve_metab, "2018-03-07 17:00:00", "2018-03-09 17:00:00"), # Mar 7-9
    mar_14 = metab_timeperiods(steve_metab, "2018-03-13 09:00:00", "2018-03-15 09:00:00"), # Mar 13-15
    mar_19 = metab_timeperiods(steve_metab, "2018-03-17 09:00:00", "2018-03-19 09:00:00"), # Mar 18-20
    mar_23 = metab_timeperiods(steve_metab, "2018-03-22 00:00:00", "2018-03-24 00:00:00"), # Mar 22-24
    mar_29 = metab_timeperiods(steve_metab, "2018-03-28 12:00:00", "2018-03-30 12:00:00"), # Mar 28-30
    apr_10 = metab_timeperiods(steve_metab, "2018-04-09 13:00:00", "2018-04-10 23:30:00"), # Apr 9-10
    apr_12 = metab_timeperiods(steve_metab, "2018-04-11 17:30:00", "2018-04-13 17:30:00"), # Apr 11-13
    apr_15 = metab_timeperiods(steve_metab, "2018-04-14 00:00:00", "2018-04-16 00:00:00"), # Apr 14-16
    apr_19 = metab_timeperiods(steve_metab, "2018-04-19 00:00:00", "2018-04-19 23:30:00") # Apr 18-20
  )
metabmodels_steve_results <- 
  data.frame(
    GPP = t(as.data.frame(lapply(metabmodels_steve, `[[`, 1))),
    ER = t(as.data.frame(lapply(metabmodels_steve, `[[`, 2)))
  )
metabmodels_steve_results$date <- strptime(row.names(metabmodels_steve_results), format = "%b_%d")
metabmodels_steve_results$date <- as.Date(metabmodels_steve_results$date)

# desoto
metabmodels_desoto <- 
  list(
    Feb_05 = metab_timeperiods(desoto_metab, "2018-02-04 00:00:00", "2018-02-06 00:00:00"), # Feb 4-6
    feb_21 = metab_timeperiods(desoto_metab, "2018-02-20 00:00:00", "2018-02-22 00:00:00"), # Feb 20-22
    feb_26 = metab_timeperiods(desoto_metab, "2018-02-25 02:00:00", "2018-02-27 02:00:00"), # Feb 25-27
    mar_03 = metab_timeperiods(desoto_metab, "2018-03-02 09:30:00", "2018-03-04 09:30:00"), # Mar 2-4
    mar_08 = metab_timeperiods(desoto_metab, "2018-03-07 17:00:00", "2018-03-09 17:00:00"), # Mar 7-9
    mar_19 = metab_timeperiods(desoto_metab, "2018-03-18 09:00:00", "2018-03-20 09:00:00"), # Mar 18-20
    mar_23 = metab_timeperiods(desoto_metab, "2018-03-22 00:00:00", "2018-03-24 00:00:00"), # Mar 22-24
    mar_30 = metab_timeperiods(desoto_metab, "2018-03-30 00:00:00", "2018-03-31 00:00:00"), # Mar 30-31
    apr_02 = metab_timeperiods(desoto_metab, "2018-04-01 00:00:00", "2018-04-03 00:00:00"), # Apr 1-3
    apr_09 = metab_timeperiods(desoto_metab, "2018-04-08 12:00:00", "2018-04-09 12:00:00"), # Apr 8-9
    apr_15 = metab_timeperiods(desoto_metab, "2018-04-14 00:00:00", "2018-04-16 00:00:00"), # Apr 14-16
    apr_19 = metab_timeperiods(desoto_metab, "2018-04-18 00:00:00", "2018-04-20 00:00:00"), # Apr 18-20
    apr_23 = metab_timeperiods(desoto_metab, "2018-04-22 00:00:00", "2018-04-24 00:00:00"), # Apr 22-24
    apr_29 = metab_timeperiods(desoto_metab, "2018-04-28 00:00:00", "2018-04-30 12:00:00") # Apr 28-30
  )
metabmodels_desoto_results <- 
  data.frame(
    GPP = t(as.data.frame(lapply(metabmodels_desoto, `[[`, 1))),
    ER = t(as.data.frame(lapply(metabmodels_desoto, `[[`, 2)))
  )
metabmodels_desoto_results$date <- strptime(row.names(metabmodels_desoto_results), format = "%b_%d")
metabmodels_desoto_results$date <- as.Date(metabmodels_desoto_results$date)
```

```{r metabPlotting, fig.width=9, fig.height=3}
GPP_plot <- 
ggplot(NULL) +
  geom_point(data = metabmodels_eric_results, aes(x = date, y = GPP, color = "Site 2",
                                                  shape = "19", size = 2), show.legend = FALSE) +
  geom_point(data = metabmodels_steve_results, aes(x = date, y = GPP, color = "Site 3", 
                                                   shape = "17", size = 2), show.legend = FALSE) +
  geom_point(data = metabmodels_desoto_results, aes(x = date, y = GPP, color = "Site 4", 
                                                    shape = "15", size = 2), show.legend = FALSE) +
  theme_classic() +
  xlab(NULL) +
  ylab("GPP (g O2 m-2 d-1)") +
  coord_cartesian(y = c(0,5), x = c(mdy("02/01/2018"), mdy("05/01/2018"))) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  scale_color_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_shape_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4]) +
    geom_vline(xintercept = mdy("04/01/2018"))
ggsave(filename = "GPP_plot.png", plot = GPP_plot, dpi = 500, width = 9, height = 3)
GPP_plot

ER_plot <- 
ggplot(NULL) +
  geom_point(data = metabmodels_eric_results, aes(x = date, y = ER, color = "Site 2",
                                                  shape = "19", size = 2), show.legend = FALSE) +
  geom_point(data = metabmodels_steve_results, aes(x = date, y = ER, color = "Site 3", 
                                                   shape = "17", size = 2), show.legend = FALSE) +
  geom_point(data = metabmodels_desoto_results, aes(x = date, y = ER, color = "Site 4", 
                                                    shape = "15", size = 2), show.legend = FALSE) +
  theme_classic() +
  xlab(NULL) +
  ylab(NULL) +
  coord_cartesian(y = c(-2,0.5), x = c(mdy("02/01/2018"), mdy("05/01/2018"))) +
  geom_vline(xintercept = mdy("04/01/2018")) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d", position = "top") +
  scale_color_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_shape_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4]) 

ER_plot
ggsave(filename = "ER.png", plot = ER_plot, dpi = 500, width = 9, height = 3)

```

# Comparison of change in N to GPP and ER

```{r comparisonofNtoMetab, fig.width=5, fig.height=5}
# GPP and ER and N data frames
PAR$date <- date(PAR$dateTime)
ericN_daynightdiff$date <- date(ericN_daynightdiff$`Date and time`)
NMetabComp_eric <- merge(ericN_daynightdiff, metabmodels_eric_results, 
                         by = "date")
NMetabComp_eric <- merge(NMetabComp_eric, PAR, by = "date")
NMetabComp_eric_summary_perc <- 
NMetabComp_eric %>%
  group_by(date) %>%
  summarise(mean_perc = mean(perc), sd_perc = sd(perc), 
            GPP = mean(GPP), ER = mean(ER), sumPAR = mean(PARMean_sum), sd_sumPAR = sd(PARMean_sum))

steveN_daynightdiff$date <- date(steveN_daynightdiff$`Date and time`)
NMetabComp_steve <- merge(steveN_daynightdiff, metabmodels_steve_results, 
                          by = "date")
NMetabComp_steve <- merge(NMetabComp_steve, PAR, by = "date")
NMetabComp_steve_summary_perc <- 
NMetabComp_steve %>%
  group_by(date) %>%
  summarise(mean_perc = mean(perc), sd_perc = sd(perc), 
            GPP = mean(GPP), ER = mean(ER), sumPAR = mean(PARMean_sum), sd_sumPAR = sd(PARMean_sum))

desotoN_daynightdiff$date <- date(desotoN_daynightdiff$dateTime)
NMetabComp_desoto <- merge(desotoN_daynightdiff, metabmodels_desoto_results,
                           by = "date")
NMetabComp_desoto <- merge(NMetabComp_desoto, PAR, by = "date")
NMetabComp_desoto_summary_perc <- 
NMetabComp_desoto %>%
  group_by(date) %>%
  summarise(mean_perc = mean(perc), sd_perc = sd(perc), 
            GPP = mean(GPP), ER = mean(ER), sumPAR = mean(PARMean_sum), sd_sumPAR = sd(PARMean_sum))

NMetabComp_eric_posperc <- NMetabComp_eric[NMetabComp_eric$perc>0,]
NMetabComp_steve_posperc <- NMetabComp_steve[NMetabComp_steve$perc>0,]
NMetabComp_desoto_posperc <- NMetabComp_desoto[NMetabComp_desoto$perc>0,]

# linear models
lm_data_GPP <- data.frame(
  GPP = c(NMetabComp_eric_posperc$GPP, NMetabComp_steve_posperc$GPP, NMetabComp_desoto_posperc$GPP),
  perc = c(NMetabComp_eric_posperc$perc, NMetabComp_steve_posperc$perc, NMetabComp_desoto_posperc$perc)
)
lm_data_ER <- data.frame(
  ER = c(NMetabComp_eric$ER, NMetabComp_steve$ER, NMetabComp_desoto$ER),
  perc = c(NMetabComp_eric$perc, NMetabComp_steve$perc, NMetabComp_desoto$perc)
)
# notation: response~terms
predicted_perc_GPP <- data.frame(
  percpred = predict.lm(lm(lm_data_GPP$perc~lm_data_GPP$GPP), lm_data_GPP), 
  GPP = lm_data_GPP$GPP
  )
predicted_perc_ER <- data.frame(
  percpred = predict.lm(lm(lm_data_ER$perc~lm_data_ER$ER), lm_data_ER), 
  ER = lm_data_ER$ER
  )

predicted_perc_GPP_eric <- data.frame(
  percpred = predict.lm(lm(NMetabComp_eric_posperc$perc~NMetabComp_eric_posperc$GPP),
                        NMetabComp_eric_posperc), 
  GPP = NMetabComp_eric_posperc$GPP
  )
predicted_perc_GPP_steve <- data.frame(
  percpred = predict.lm(lm(NMetabComp_steve_posperc$perc~NMetabComp_steve_posperc$GPP),
                        NMetabComp_steve_posperc), 
  GPP = NMetabComp_steve_posperc$GPP
  )
predicted_perc_GPP_desoto <- data.frame(
  percpred = predict.lm(lm(NMetabComp_desoto_posperc$perc~NMetabComp_desoto_posperc$GPP),
                        NMetabComp_desoto_posperc), 
  GPP = NMetabComp_desoto_posperc$GPP
  )
# GPP plot
ggplot(NULL) +
  geom_point(data = NMetabComp_eric_posperc, 
             aes(x = GPP, y = perc, color = "Site 2", shape = "19", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_steve_posperc, 
             aes(x = GPP, y = perc, color = "Site 3", shape = "17", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_desoto_posperc, 
             aes(x = GPP, y = perc, color = "Site 4", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_line(data = predicted_perc_GPP_eric, aes(y = percpred, x = GPP, color = "Site 2", size = 1.2)) +
  geom_line(data = predicted_perc_GPP_steve, aes(y = percpred, x = GPP, color = "Site 3", size = 1.2)) +
  geom_line(data = predicted_perc_GPP_desoto, aes(y = percpred, x = GPP, color = "Site 4", size = 1.2)) +
  theme_classic() +
  xlab("GPP (g O2 m-2 d-1)") +
  ylab("Change in NOX %") +
  coord_cartesian(x = c(0,5), y = c(0,2)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "none") +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])

           # GPP: pre and post 1 Apr
NMetabComp_eric_before1Apr <- NMetabComp_eric_posperc[NMetabComp_eric_posperc$date<mdy("04-01-2018"),]
NMetabComp_steve_before1Apr <- NMetabComp_steve_posperc[NMetabComp_steve_posperc$date<mdy("04-01-2018"),]
NMetabComp_desoto_before1Apr <- NMetabComp_desoto_posperc[NMetabComp_desoto_posperc$date<mdy("04-01-2018"),]

NMetabComp_eric_after1Apr <- NMetabComp_eric_posperc[NMetabComp_eric_posperc$date>mdy("04-01-2018"),]
NMetabComp_steve_after1Apr <- NMetabComp_steve_posperc[NMetabComp_steve_posperc$date>mdy("04-01-2018"),]
NMetabComp_desoto_after1Apr <- NMetabComp_desoto_posperc[NMetabComp_desoto_posperc$date>mdy_hms("04-01-2018"),]

lm_data_GPP_duringdisch <- data.frame(
  GPP = c(NMetabComp_eric_before1Apr$GPP, 
          NMetabComp_steve_before1Apr$GPP, 
          NMetabComp_desoto_before1Apr$GPP),
  perc = c(NMetabComp_eric_before1Apr$perc, 
           NMetabComp_steve_before1Apr$perc, 
           NMetabComp_desoto_before1Apr$perc)
)
predicted_perc_GPP_duringdisch <- data.frame(
  percpred = predict.lm(lm(perc~GPP, data = lm_data_GPP_duringdisch),
                        lm_data_GPP_duringdisch), 
  GPP = lm_data_GPP_duringdisch$GPP
  )

lm_data_GPP_afterdisch <- data.frame(
  GPP = c(NMetabComp_eric_after1Apr$GPP, 
          NMetabComp_steve_after1Apr$GPP, 
          NMetabComp_desoto_after1Apr$GPP),
  perc = c(NMetabComp_eric_after1Apr$perc, 
           NMetabComp_steve_after1Apr$perc, 
           NMetabComp_desoto_after1Apr$perc)
)
predicted_perc_GPP_afterdisch <- data.frame(
  percpred = predict.lm(lm(perc~GPP, data = lm_data_GPP_afterdisch),
                        lm_data_GPP_afterdisch), 
  GPP = lm_data_GPP_afterdisch$GPP
  )

ggplot(NULL) +
  geom_point(data = NMetabComp_eric_before1Apr, 
             aes(x = GPP, y = perc, color = "before", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_steve_before1Apr, 
             aes(x = GPP, y = perc, color = "before", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_desoto_before1Apr, 
             aes(x = GPP, y = perc, color = "before", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_eric_after1Apr, 
             aes(x = GPP, y = perc, color = "after", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_steve_after1Apr, 
             aes(x = GPP, y = perc, color = "after", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_desoto_after1Apr, 
             aes(x = GPP, y = perc, color = "after", shape = "15", size = 2), 
             show.legend = FALSE) +
  geom_line(data = predicted_perc_GPP_duringdisch, aes(y = percpred, x = GPP, color = "before")) +
  geom_line(data = predicted_perc_GPP_afterdisch, aes(y = percpred, x = GPP, color = "after")) +
  theme_classic() +
  xlab("GPP (g O2 m-2 d-1)") +
  ylab("Change in NOX %") +
  coord_cartesian(x = c(0,5),y = c(0,2.0)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "none") +
  scale_color_manual(values = c("green", "dimgrey"))



# ER
ggplot(NULL) +
  geom_point(data = NMetabComp_eric_summary_perc, 
             aes(x = ER, y = mean_perc, color = "Site 2",
                 shape = "19", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_eric_summary_perc, 
                aes(x = ER, ymin = mean_perc-sd_perc, ymax = mean_perc+sd_perc)) +
  geom_point(data = NMetabComp_steve_summary_perc, 
             aes(x = ER, y = mean_perc, color = "Site 3",
                 shape = "17", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_steve_summary_perc, 
                aes(x = ER, ymin = mean_perc-sd_perc, ymax = mean_perc+sd_perc)) +
  geom_point(data = NMetabComp_desoto_summary_perc, 
             aes(x = ER, y = mean_perc, color = "Site 4",
                 shape = "15", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_desoto_summary_perc, 
                aes(x = ER, ymin = mean_perc-sd_perc, ymax = mean_perc+sd_perc)) +
  geom_line(data = predicted_perc_ER, aes(y = percpred, x = ER)) +
  theme_classic() +
  xlab("ER (g O2 m-2 d-1)") +
  ylab("Change in NOX %") +
  scale_y_continuous(position = "right") +
  #coord_cartesian(x = c(-5,1), y = c(-2,2)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_color_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_shape_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])



```



```{r Metab and temp, fig.width=5, fig.height=5}
eric_metab$date <- date(eric_metab$dateTime)
NMetabComp_eric <- merge(NMetabComp_eric, eric_metab, by = "date")
NMetabComp_eric_summary <- 
NMetabComp_eric %>%
  group_by(date) %>%
  summarise(mean_temp = mean(temp), sd_temp = sd(temp), GPP = mean(GPP), ER = mean(ER))

steve_metab$date <- date(steve_metab$dateTime)
NMetabComp_steve <- merge(NMetabComp_steve, steve_metab, by = "date")
NMetabComp_steve_summary <- 
NMetabComp_steve %>%
  group_by(date) %>%
  summarise(mean_temp = mean(temp), sd_temp = sd(temp), GPP = mean(GPP), ER = mean(ER))

desoto_metab$date <- date(desoto_metab$dateTime)
NMetabComp_desoto <- merge(NMetabComp_desoto, desoto_metab, by = "date")
NMetabComp_desoto_summary <- 
NMetabComp_desoto %>%
  group_by(date) %>%
  summarise(mean_temp = mean(temp), sd_temp = sd(temp), GPP = mean(GPP), ER = mean(ER))

# linear models
lm_data_GPP_temp <- data.frame(
  GPP = c(NMetabComp_eric_summary$GPP, 
          NMetabComp_steve_summary$GPP, 
          NMetabComp_desoto_summary$GPP),
  temp = c(NMetabComp_eric_summary$mean_temp, 
           NMetabComp_steve_summary$mean_temp, 
           NMetabComp_desoto_summary$mean_temp)
)
lm_data_ER_temp <- data.frame(
  ER = c(NMetabComp_eric_summary$ER, 
         NMetabComp_steve_summary$ER, 
         NMetabComp_desoto_summary$ER),
  temp = c(NMetabComp_eric_summary$mean_temp, 
           NMetabComp_steve_summary$mean_temp, 
           NMetabComp_desoto_summary$mean_temp)
)
# notation: response~terms
predicted_temp_GPP <- data.frame(
  GPPpred = predict.lm(lm(lm_data_GPP_temp$GPP~lm_data_GPP_temp$temp), lm_data_GPP_temp), 
  temp = lm_data_GPP_temp$temp
  )
predicted_temp_ER <- data.frame(
  temppred = predict.lm(lm(lm_data_ER_temp$temp~lm_data_ER_temp$ER), lm_data_ER_temp), 
  ER = lm_data_ER_temp$ER
  )

# GPP and temp plot
ggplot(NULL) +
  geom_point(data = NMetabComp_eric_summary, 
             aes(y = GPP, x = mean_temp, color = "Site 2", shape = "19", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_eric_summary, 
                #aes(y = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_steve_summary, 
             aes(y = GPP, x = mean_temp, color = "Site 3", shape = "17", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_steve_summary, 
                #aes(x = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_desoto_summary, 
             aes(y = GPP, x = mean_temp, color = "Site 4", shape = "15", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_desoto_summary, 
                #aes(x = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_line(data = predicted_temp_GPP, aes(x = temp, y = GPPpred)) +
  theme_classic() +
  xlab("temp") +
  ylab("gpp") +
  coord_cartesian(y = c(0,5), x = c(0,20)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])

# ER and temp plot
ggplot(NULL) +
  geom_point(data = NMetabComp_eric_summary, aes(x = ER, y = mean_temp, color = "Site 2",
                                           shape = "19", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_eric_summary, 
                aes(x = ER, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_steve_summary, aes(x = ER, y = mean_temp, color = "Site 3",
                                           shape = "17", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_steve_summary, 
                aes(x = ER, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_desoto_summary, aes(x = ER, y = mean_temp, color = "Site 4",
                                           shape = "15", size = 2), show.legend = FALSE) +
  geom_errorbar(data = NMetabComp_desoto_summary, 
                aes(x = ER, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_line(data = predicted_temp_ER, aes(y = temppred, x = ER)) +
  theme_classic() +
  xlab("ER (g O2 m-2 d-1)") +
  ylab("mean temp") +
  scale_y_continuous(position = "right") +
  #coord_cartesian(x = c(-5,1), y = c(-2,2)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_color_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_shape_discrete(name = "Site", breaks = c("Site 2", "Site 3", "Site 4"), 
                       labels = c("Site 2", "Site 3", "Site 4")) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])
```

```{r GPPandpar, fig.width=5, fig.height=5}
lm_data_GPP_par <- data.frame(
  GPP = c(NMetabComp_eric_summary$GPP, 
          NMetabComp_steve_summary$GPP, 
          NMetabComp_desoto_summary$GPP),
  par = c(NMetabComp_eric_summary_perc$sumPAR, 
           NMetabComp_steve_summary_perc$sumPAR, 
           NMetabComp_desoto_summary_perc$sumPAR)
)
predicted_par_GPP <- data.frame(
  GPPpred = predict.lm(lm(lm_data_GPP_par$GPP~lm_data_GPP_par$par), lm_data_GPP_par), 
  par = lm_data_GPP_par$par*1.0E-6
  )

# par units: mol m-2 s-1
ggplot(NULL) +
  geom_point(data = NMetabComp_eric_summary_perc, 
             aes(y = GPP, x = sumPAR*1.0E-6, color = "Site 2", shape = "19", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_eric_summary_perc, 
                #aes(y = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_steve_summary_perc, 
             aes(y = GPP, x = sumPAR*1.0E-6, color = "Site 3", shape = "17", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_steve_summary, 
                #aes(x = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_point(data = NMetabComp_desoto_summary_perc, 
             aes(y = GPP, x = sumPAR*1.0E-6, color = "Site 4", shape = "15", size = 2), 
             show.legend = FALSE) +
  #geom_errorbar(data = NMetabComp_desoto_summary, 
                #aes(x = GPP, ymin = mean_temp-sd_temp, ymax = mean_temp+sd_temp)) +
  geom_line(data = predicted_par_GPP, aes(x = par, y = GPPpred)) +
  theme_classic() +
  xlab("PAR") +
  ylab("gpp") +
  coord_cartesian(y = c(0,5), x = c(0,0.5)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])
```

```{r NtoMetabbyTime, fig.width=5, fig.height=5}
# GPP plot: before 1 Apr
NMetabComp_eric_before1Apr <- NMetabComp_eric[NMetabComp_eric$date<mdy("04-01-2018"),]
NMetabComp_steve_before1Apr <- NMetabComp_steve[NMetabComp_steve$date<mdy("04-01-2018"),]
NMetabComp_desoto_before1Apr <- NMetabComp_desoto[NMetabComp_desoto$date<mdy_hms("04-01-2018"),]

ggplot(NULL) +
  geom_point(data = NMetabComp_eric_before1Apr, 
             aes(x = GPP, y = perc, color = "Site 2", shape = "19", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_steve_before1Apr, 
             aes(x = GPP, y = perc, color = "Site 3", shape = "17", size = 2), 
             show.legend = FALSE) +
  geom_point(data = NMetabComp_desoto_before1Apr, 
             aes(x = GPP, y = perc, color = "Site 4", shape = "15", size = 2), 
             show.legend = FALSE) +
  #geom_line(data = predicted_perc_GPP, aes(y = percpred, x = GPP)) +
  theme_classic() +
  xlab("GPP (g O2 m-2 d-1)") +
  ylab("Change in NOX %") +
  coord_cartesian(y = c(0,1.5)) +
  ggtitle(NULL) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_color_manual(breaks = c("Site 2", "Site 3", "Site 4"),
                     values = colorBlindPalette[2:4])


```

```{r 24Hr and NO3 GPP, fig.width=5, fig.height=5}
#30 mar, 9 apr
unique(NMetabComp_eric_posperc$date) #29 mar, 9 apr
unique(NMetabComp_steve_posperc$date) # 29 Mar, 10 Apr
unique(NMetabComp_desoto_posperc$date) #29 mar, 2 apr

hr24 <- 
data.frame(
  time = c("March", "March", "March", "April", "April", "April"),
  location = c("2", "3", "4", "2", "3", "4"),
  perc = c(
    na.omit(unique(NMetabComp_eric_posperc$perc[NMetabComp_eric_posperc$date == ymd("2018-03-29")])),
    na.omit(unique(NMetabComp_steve_posperc$perc[NMetabComp_steve_posperc$date == ymd("2018-03-29")])),
    na.omit(unique(NMetabComp_desoto_posperc$perc[NMetabComp_desoto_posperc$date == ymd("2018-04-02")])),
    na.omit(unique(NMetabComp_eric_posperc$perc[NMetabComp_eric_posperc$date == ymd("2018-04-09")])),
    na.omit(unique(NMetabComp_steve_posperc$perc[NMetabComp_steve_posperc$date == ymd("2018-04-10")])),
    na.omit(unique(NMetabComp_desoto_posperc$perc[NMetabComp_desoto_posperc$date == ymd("2018-04-09")]))[2]
  ),
  GPP = c(
    na.omit(unique(NMetabComp_eric_posperc$GPP[NMetabComp_eric_posperc$date == ymd("2018-03-29")])),
    na.omit(unique(NMetabComp_steve_posperc$GPP[NMetabComp_steve_posperc$date == ymd("2018-03-29")])),
    na.omit(unique(NMetabComp_desoto_posperc$GPP[NMetabComp_desoto_posperc$date == ymd("2018-04-02")])),
    na.omit(unique(NMetabComp_eric_posperc$GPP[NMetabComp_eric_posperc$date == ymd("2018-04-09")])),
    na.omit(unique(NMetabComp_steve_posperc$GPP[NMetabComp_steve_posperc$date == ymd("2018-04-10")])),
    na.omit(unique(NMetabComp_desoto_posperc$GPP[NMetabComp_desoto_posperc$date == ymd("2018-04-09")]))
  )
)

ggplot(NULL) +
  geom_point(data = hr24_March, aes(x = GPP, y = perc, color = "March", shape = "19", size = 2)) +
  geom_point(data = hr24_Apr, aes(x = GPP, y = perc, color = "Apr", shape = "19", size = 2))


ggplot(data = hr24, aes(fill = location)) +
  geom_bar(aes(x = time, y = perc)) +
  geom_point(aes(x = location, y = GPP/10, size = 2)) +
  scale_y_continuous(sec.axis = sec_axis(~./10)) +
  ggtitle(NULL) +
    theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
    scale_color_manual(breaks = location,
                     values = colorBlindPalette[2:4]) +
  theme_classic()

# this one
ggplot(NULL) + 
    geom_col(aes(fill = location, x = time, y = perc), data = hr24, position = "dodge") +
    geom_boxplot(aes(fill = location, x = time, y = GPP), data = hr24) +
    ggtitle(NULL) +
    scale_color_manual(breaks = location,
                     values = colorBlindPalette[2:4]) +
    scale_x_discrete(limits=c("March", "April")) +
    theme_classic() +
    theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
    ylab("percent change (bars) or GPP (lines)") +
    scale_color_manual(breaks = c(2,3,4), values = colorBlindPalette[2:4])

ggplot(data = hr24, aes(x = time, y = GPP)) + 
    geom_boxplot(aes(fill = location), position = "dodge") +
    ggtitle(NULL) +
    theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
    scale_color_manual(breaks = location,
                     values = colorBlindPalette[2:4]) +
  theme_classic()

# red is march, blue is april
ggplot(NULL) +
  geom_point(data = hr24, aes(x = GPP, y = perc, color = time, shape = location, size = 3)) +
  theme_classic() +
  coord_cartesian(xlim = c(0,4), ylim = c(0,0.6)) +
  scale_shape_manual(breaks = c("2", "3", "4"), values = c(15,17,19)) +
  scale_color_manual(breaks = c("March, April"), values = c("blue", "red")) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        legend.position = "none")
        


```
